---
title: "Exploring rugby data"
author: "Jocelyn Mara"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


## Overview

This document describes the exploratory analysis conducted on data related to try-scoring in the 2017 Super Rugby competition. 

This data consists of tries that were scored during the 2017 Super Rugby competition (observations/rows). Here is a description of the variables:

1. `try_no`: a unique identification number given to each try
2. `round_no`: an identification number to distinguish the round number the try was scored in
3. `attacking_team`: the try-scoring team
4. `defending_team`: the opposition team who conceded the try
5. `attacking_rank`: the final league ranking at the end of the season of the try-scoring team
6. `defending_rank`: the final league ranking at the end of the season of the opposition team
7. `attacking_conference`: the conference group of the try-scoring team
8. `defending_conference`: the conference group of the opposition team
9. `game_time`: the game time in minutes when the try was scored
10. `try_source`: the initial source of possession for the attacking team preceding the try
11. `final_source`: the event that directly preceded the try and resulted in the try being scored
12. `phases`: the total number of phases between gaining possession, and the try being scored (a phase is from one ruck to the next ruck)
13. `time_from_source`: the time taken from gaining possession to scoring the try, in seconds
14. `possession_zone`: the zone on the field the attacking team gained possession of the ball before scoring the try (A = attacking 22m line to try-line, B = halfway to attacking 22m line, C = defensive 22m line to halfway, D = )
15. `offloads`: the number of offloads from gaining possession to the try being scored
15. `passes`: the number of passes from gaining possession to the try being scored
16. `total_passes`: the number of offloads plus passes 

This data was collected by a former University of Canberra student, Molly Coughlan, as part of a project that identified playing patterns that led to tries in super rugby^[Coughlan, Mountifield, Sharpe & Mara, 2019. How they scored the tries: applying cluster analysis to identify playing patterns that lead to tries in super rugby. IJPAS.]

## Packages

The following packages will be loaded and used in this analysis:

```{r}
library(tidyverse)
library(naniar)
```


## Reading in the data

The data can be read into RStudio and we can examine the structure using the following:

```{r}
df <- read_csv("data/2017_super-rugby_try-source-data.csv")
str(df)
```


## Checking for missing values

You can check for missing values by using a visualisation such as the `vis_miss()` function from the `naniar` package:

```{r}
vis_miss(df)
```


Alternatively, you can check how many missing values there are using the following: 

```{r}
sum(is.na(df))
```

You can also check which rows and columns the missing values are in using:

```{r, eval = FALSE}
which(is.na(df), arr.ind = TRUE)
# output not printed here as too long
```


## Dealing with missing values

The missing values in this data actually represent 0's. For example, if there were no passes for a try, that was left blank.

We can replace the missing values with 0's using the following:

```{r}
# if there are multiple variables that contain missing values, 
# you need to state what to replace NAs with for each one

df <- replace_na(data = df, replace = list(phases = 0, 
                                           offloads = 0, 
                                           passes = 0, 
                                           total_passes = 0))
```

Let's double check that has done the job:

```{r}
sum(is.na(df))
```


## Exploratory visualisations



## References